---
title: "Your A4 Report Title (TODO: Revise title)"
author: "Molly Banks"
date: '2022-11-09'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/analysis.R")
```

## Abstract 
[TODO: Include a short abstract (about 50 words)]


## Introduction 
[TODO]

## Data Summary 
[TODO]

## Section 3: Growth of the U.S. Prison Population

```{r prison, echo=FALSE, message=FALSE, warning=FALSE}
plot_jail_pop_for_us()
```

Summary Paragraph 


## Section 4: Growth of Prison Population by State
```{r state, echo=FALSE, message=FALSE, warning=FALSE}
plot_jail_pop_by_states(c("WA", "NE", "AL", "NJ"))
```

summary Paragraph 
- selected one state from each region specified in the Vera institute data set

## Section 5: Pretrial Jailing Rate Inequality by Urbanicity
```{r pretrial, echo=FALSE, message=FALSE, warning=FALSE}
inequality_plot()
```

summary Paragraph 

## Section 6: Pretrial Jailing Rate Inequality by State
```{r map, echo=FALSE, message=FALSE, warning=FALSE}
plot_county_ineq("WA", "2010")
```

summary Paragraph 


