---
title: "Injustice in the Justice System: Noteworthy US Incarceration Trends 1970-2018"
author: "Molly Banks"
date: '2022-11-09'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/analysis.R")
```

## Abstract 

Incarceration trends since 1970  reveal stark inequalities in the American justice system. While policy in the justice system has changed throughout the decades, regions, states, and counties have felt starkly different effects. While incarceration remains incredibly high, this report examines historical incarceration trends and potential inequalities that might illuminate future solutions. 

## Introduction 

 

## Data Summary 
In  `r year_max_pretrial_rate`, `r county_max_pretrial_rate`'s pretrial jailing rate was the highest in America between 1970 and 2018 within all of America. 
In `r year_max_ice_jail`, `r county_max_ice_jail`, TX had the highest number of jail admits from ICE. Examining jail bookings by location reveals particular locations of interest in evaluating American incarceration trends. 

## Section 3: Growth of the U.S. Prison Population

```{r prison, echo=FALSE, message=FALSE, warning=FALSE}
plot_jail_pop_for_us()
```

How has America’s prison population grown in the aftermath of the “tough on crime” era? This graph depicts a significant increase in the prison population since the mid-1970s. There is a clear increase that begins that the start of the “tough on crime” era. From 1990 to 2010, incarceration rates skyrocketed. 


## Section 4: Growth of Prison Population by State
```{r state, echo=FALSE, message=FALSE, warning=FALSE}
plot_jail_pop_by_states(c("WA", "NE", "AL", "NJ"))
```

The selected four states represent the four US regions specified by the Vera Project. Each state had distinct trends in prison population growth.  This suggests that the “tough on crime” era functioned differently at the state level. New Jersey’s new incarcerations skyrocketed at the height of this era but, along with Alabama, declined in the aftermath. Nebraska and Washington, on the other hand, are still steadily increasing in total prison populations. 

## Section 5: Pretrial Jailing Rate Inequality by Urbanicity
```{r pretrial, echo=FALSE, message=FALSE, warning=FALSE}
inequality_plot()
```

How did rural and urban counties differ in pretrial jailing rates after the turn of the century? While urban pretrial jailing has decreased since the 2000s, pretrial jailing in rural counties has continued to increase. The policy changes over the last two decades have had distinct impacts depending on urbanicity. 

## Section 6: Pretrial Jail Bookings by State
```{r map, echo=FALSE, message=FALSE, warning=FALSE}
plot_county_ineq("WA", "2007")
```

A speedy trial requires resources. In Washington state, rural states have particularly high pretrial jailing rates. This map suggests that rural counties lack the infrastructure necessary to support an effective justice system. In 2007, the aftermath of the “tough on crime” was still roaring on. This map illuminates how the aftermath played out at the county level.


